Supplementary Materials for ``Mott Criticality as the Confinement Transition of a Pseudogap-Mott Metal"

### 2025-12-10 submissions/Supplementary/supplementary.tex
started creating referee response

%\documentclass{article}
%\documentclass[
%%preprint,
%reprint,onecolumn,
%superscriptaddress,
%groupedaddress,
%%unsortedaddress,
%%runinaddress,
%%frontmatterverbose, 
%%preprint,
%%preprintnumbers,
%%nofootinbib,
%%nobibnotes,
%%bibnotes,
%%aps,superscriptaddress,
%aps,pra,
%%rmp,
%%prstab,
%%prstper,
%%floatfix,
%10pt
%]{revtex4-2}
\documentclass[%
 reprint,onecolumn,
superscriptaddress,
groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
aps,superscriptaddress,
pra,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}
%\usepackage{amsmath,amssymb,bm,graphicx,braket}
\usepackage[margin=1.5cm]{geometry}
%\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{graphicx,braket,color,bm,amsmath,amssymb,braket,placeins}
% \graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}
\usepackage{natbib}

\usepackage{etoolbox}
\renewcommand{\thefigure}{S\arabic{figure}}
\makeatletter
\apptocmd{\thebibliography}{\global\c@NAT@ctr 89\relax}{}{}
\makeatother
\begin{document}
\title{Supplementary Materials for ``Mott Criticality as the Confinement Transition of a Pseudogap-Mott Metal"}

\author{Abhirup Mukherjee}
\email{am18ip014@iiserkol.ac.in}
\affiliation{Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}
\author{S. R. Hassan}
\email{shassan@imsc.res.in}
\affiliation{The Institute of Mathematical Sciences, HBNI, C.I.T. Campus, Chennai 600 113, India}
\author{Anamitra Mukherjee}
\email{anamitra@niser.ac.in}
\affiliation{School of Physical Sciences, National Institute of Science, Education and Research, HBNI, Jatni 752050, India}
\author{N. S. Vidhyadhiraja}
\email{raja@jncasr.ac.in}
\affiliation{Theoretical Sciences Unit, Jawaharlal Nehru Center for Advanced Scientific Research, Jakkur, Bengaluru 560064, India}
\author{A. Taraphder}
\email{arghya@phy.iitkgp.ac.in}
\affiliation{Department of Physics, Indian Institute of Technology Kharagpur, Kharagpur 721302, India}
\author{Siddhartha Lal}
\email{slal@iiserkol.ac.in}
\affiliation{Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}

@@

The eigenstates \(\ket{\Psi_{s}}\) (\(s = \left({\bf k}, n\right)\)) of the lattice Hamiltonians 
%obtained using eq.~\ref{tilingPrescriptionFinal} also 
enjoy a {\it many-body} Bloch's theorem~[55]
%\cite{stoyanova}, 
because the tiling procedure restores the translation symmetry of the Hamiltonian (as shown in eq.~\ref{translationSymmetry}). This means that the {\it local} eigenstates \(\ket{\psi_n\left({\bf r}_d\right)}\) (with the impurity located at an arbitrary position \({\bf r}_d\)) of the unit cell auxiliary model Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) %defined in eq.~\ref{unitCellHamiltonian} 
can be used to construct eigenstates of the lattice Hamiltonian. The index \(n(=0,1,\ldots)\) in the subscript indicates that it is the \(n^\text{th}\) eigenstate of the auxiliary model.

The state \(\ket{\psi_n\left({\bf r}_d\right)}\) does not specify the position of the zeroth site, because the unit cell Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) itself has been averaged over \(\mathcal{Z}\) zeroth sites. Accordingly, we can express the averaged eigenstate \(\ket{\psi_n\left({\bf r}_d\right)}\) as
\begin{equation}\begin{aligned}
	\ket{\psi_n\left({\bf r}_d\right)} = \frac{1}{\sqrt\mathcal{Z}}\sum_{{\bf z} \in \text{NN}({\bf r}_d)}\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}\) is an auxiliary model eigenstate with the impurity and zeroth sites placed at \({\bf r}_d\) and \({\bf z}\). With this in mind, the following unnormalised combination of the auxiliary model eigenstates satisfies a many-particle equivalent of Bloch's theorem~[55]:
%\cite{stoyanova}:
\begin{equation}\begin{aligned}\label{eigenstateProposal}
	\ket{\Psi_{s}} \equiv \ket{\Psi_{{\bf k}, n}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(N\) is the total number of lattice sites and \({\bf r}_d\) is summed over all lattice spacings. The set of \(n=0\) states form the lowest band in the spectrum of the lattice, while higher values of \(n\) produce the more energetic bands. The ground state \(s = s_0\) is obtained by setting \({\bf k}\) and \(n\) to 0:
\begin{equation}\begin{aligned}\label{groundstateProposal}
	\ket{\Psi_\text{gs}} \equiv \ket{\Psi_{s_0}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d, {\bf z}\right)}
\end{aligned}\end{equation}

@@

\subsection{The unitary renormalisation group method}
In order to obtain the various low-energy phases of our impurity model, we perform a scaling analysis of the associated Hamiltonian using the recently developed unitary renormalisation group (URG) method ~[24,42,43].
%\cite{anirbanurg1,anirbanurg2}. 
The method has been applied successfully on a wide variety of problems of correlated fermions~\cite{santanukagome,1dhubjhep1,anirbanmott1,siddharthacpi,anirban_kondo,Patra_2023}. The method proceeds by resolving quantum fluctuations in high-energy degrees of freedom, leading to a low-energy Hamiltonian with renormalised couplings and new emergent degrees of freedom. Typically, for a system with Fermi energy \(\epsilon_F\) and bandwidth \(E_N\), the sequence of isoenergetic shells \(\left\{E_{(j)}\right\}, E_{j}\in \left[E_0, E_N\right] \) define the states whose quantum fluctuations we sequentially resolve. The momentum states lying on shells \(E_N\) that are far away from the Fermi surface comprise the UV states, while those on shells near the Fermi surface comprise the IR states.

As a result of the URG transformations, the Hamiltonian \(H_{(j)}\) at a given RG step \(j\) involves scattering processes between the \(k-\)states that have energies lower than \(D_{(j+1)}\). The unitary transformation \(U_{(j)}\) is then defined so as to remove the number fluctuations of the currently most energetic set of states \(D_{(j)}\)~[42,43]:
%\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	H_{(j-1)} = U_{(j)} H_{(j)} U^\dagger_{(j)}~, \text{such that} ~\left[H_{(j-1)}, \hat n_{j}\right] =0~.
\end{eqnarray}
The eigenvalue of $\hat{n}_{j}$ has, thus, been rendered an integral of motion (IOM) under the RG transformation.

The unitary transformations can be expressed in terms of a generator \(\eta_{(j)}\) that has fermionic algebra~[42,43]:
%\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	\label{unitary}
	U_{(j)} = \frac{1}{\sqrt 2}\left(1 + \eta_{(j)} - \eta_{(j)}^\dagger\right)~,~ \quad\left\{ \eta_{(j)},\eta_{(j)}^\dagger \right\} = 1~,
\end{eqnarray}
where \(\left\{\cdot\right\}\) is the anticommutator. The unitary operator \(U_{(j)}\) that appears in Eq.~\eqref{unitary} can be cast into the well-known general form \(U = e^\mathcal{S}, \mathcal{S} = \frac{\pi}{4}\left( \eta^\dagger_{(j)} - \eta_{(j)} \right)\) that a unitary operator can take, defined by an anti-Hermitian operator \(\mathcal{S}\). The generator \(\eta_{(j)}\) is given by the expression~[42,43]:
%\cite{anirbanurg1,anirbanurg2}
\begin{eqnarray}
	\eta^\dagger_{(j)} = \frac{1}{\hat \omega_{(j)} - \text{Tr}\left(H_{(j)} \hat n_{j}\right) } c^\dagger_{j} \text{Tr}\left(H_{(j)}c_{j}\right)~.
\end{eqnarray}
The operators \(\eta_{(j)},\eta^\dagger_{(j)}\) behave as the many-particle analogues of the single-particle field operators \(c_j,c^\dagger_j\) - they change the occupation number of the single-particle Fock space \(\ket{n_j}\).  The important operator \(\hat \omega_{(j)}\) originates from the quantum fluctuations that exist in the problem because of the non-commutation of the kinetic energy terms and the interaction terms in the Hamiltonian:
\begin{eqnarray}
	\hat \omega_{(j)} = H_{(j-1)} - H^i_{(j)}~.
	\label{omega}
\end{eqnarray}
\(H^i_{(j)}\) is the part of \(H_{(j)}\) that commutes with \(\hat n_j\) but does {\it not} commute with at least one \(\hat n_l\) for \(l < j\). The RG flow continues up to energy \(D^*\), where a fixed point is reached from the vanishing of the RG function. 
Detailed comparisons of the URG with other methods (e.g., the functional RG, spectrum bifurcation RG etc.) can be found in Refs.~[42,43].
%\cite{anirbanmott1,anirbanurg1}.

@@

\section{Tiling formalism: Mapping from the auxiliary model to the tiled lattice model}
The eigenstates \(\ket{\Psi_{s}}\) (\(s = \left({\bf k}, n\right)\)) of the lattice Hamiltonians obtained using  the tiling procedure outlined in the main text enjoy a {\it many-body} Bloch's theorem~[55],
%\cite{stoyanova}, 
because the tiling procedure restores the translation symmetry of the Hamiltonian. This means that the {\it local} eigenstates \(\ket{\psi_n\left({\bf r}_d\right)}\) (with the impurity located at an arbitrary position \({\bf r}_d\)) of the unit cell auxiliary model Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) can be used to construct eigenstates of the lattice Hamiltonian. The index \(n(=0,1,\ldots)\) in the subscript indicates that it is the \(n^\text{th}\) eigenstate of the auxiliary model.

The state \(\ket{\psi_n\left({\bf r}_d\right)}\) does not specify the position of the zeroth site, because the unit cell Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) itself has been averaged over \(\mathcal{Z}\) zeroth sites. Accordingly, we can express the averaged eigenstate \(\ket{\psi_n\left({\bf r}_d\right)}\) as
\begin{equation}\begin{aligned}
	\ket{\psi_n\left({\bf r}_d\right)} = \frac{1}{\sqrt\mathcal{Z}}\sum_{{\bf z} \in \text{NN}({\bf r}_d)}\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}\) is an auxiliary model eigenstate with the impurity and zeroth sites placed at \({\bf r}_d\) and \({\bf z}\). With this in mind, the following unnormalised combination of the auxiliary model eigenstates satisfies a many-particle equivalent of Bloch's theorem~[55]:
%\cite{stoyanova}:
\begin{equation}\begin{aligned}\label{eigenstateProposal1}
	\ket{\Psi_{s}} \equiv \ket{\Psi_{{\bf k}, n}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(N\) is the total number of lattice sites and \({\bf r}_d\) is summed over all lattice spacings. The set of \(n=0\) states form the lowest band in the spectrum of the lattice, while higher values of \(n\) produce the more energetic bands. The ground state \(s = s_0\) is obtained by setting \({\bf k}\) and \(n\) to 0:
\begin{equation}\begin{aligned}\label{groundstateProposal1}
	\ket{\Psi_\text{gs}} \equiv \ket{\Psi_{s_0}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d, {\bf z}\right)}
\end{aligned}\end{equation}

@@

The ${\bf q}_1={\bf q}_2$ component of the Hamiltonian shows the emergence of the exactly solvable Hatsugai-Kohmoto model~[45,46]
%\cite{Baskaran1991,Hatsugai1992} 
at the critical point. The correlation term $\mathcal{U}\sum_{{\bf q}, \sigma}r_{{\bf q} \sigma}r_{{\bf q} \bar\sigma}$ leads to a transfer of spectral weight across the Fermi surface and separates the available states into three classes:
\begin{equation}\begin{aligned}
\braket{n_{\bf q}} = \begin{cases}
    2, \quad \epsilon_{\bf q} < -|\mathcal{U}/2|~,\\
    1, \quad |\mathcal{U}/2| > \epsilon_{\bf q} > -|\mathcal{U}/2|~,\\
    0, \quad |\mathcal{U}/2| < \epsilon_{\bf q}~.
\end{cases}
\end{aligned}\end{equation}
Different from a Fermi liquid is the emergence of the highly degenerate single-occupied region in the middle, and this gives rise to non-Fermi liquid excitations.

@@

\section{Universal scaling of impurity spectral function}
In the main manuscript, we demonstrate how the PG phase is characterised by a universal scaling exponent \(\beta\) for the self-energy: \(-1/\Sigma^{\prime\prime}(\omega)\sim 1/\Sigma^{\prime\prime}(0) + \omega^{\beta},~\beta=2\). We have verified that a similar behaviour holds for the low-frequency regime of the impurity spectral function (see Fig.~\ref{optical}, left panel). We observe that $A_{d}(\omega)$ scales as $A_d(0) + \omega^{2}$ for the pseudogap phase, while $1/A_{d}(\omega)\sim 1/A_{d}(0) + \omega^{2}$ for the Fermi liquid (inset of Fig.~\ref{optical}).

\section{Finite-Frequency Optical Response Through the Pseudogap}
\begin{figure}[!ht]
    \includegraphics[width=0.45\linewidth]{Ad_fit.pdf}
	\hfill
	\includegraphics[width=0.45\textwidth]{sigma_in.pdf}
    \caption{Left: Impurity spectral function $A_d(\omega)$ in the local Fermi liquid (inset) and pseudogap (main). The PG spectral function is fit to a power law (with the $\omega=0$ contribution subtracted, see figure in the main manuscript for the full spectral function). The exponent of the power law comes out to be $n\approx 2$ throughout the PG. The local FL spectral function (inset) is fit to a lorentzian (by fitting the inverse to a power law).\\ Right: Optical conductivity $\sigma(\omega)$ for excitations of the impurity model, approximated through the product of the carrier density (taken from the density of states $A(\omega)$) and the lifetime $\tau(\omega)$ (obtained from the inverse scattering rate $-1/\Sigma^{\prime\prime}(\omega)$, where $\Sigma^{\prime\prime}$ is the imaginary part of the self-energy). The Fermi liquid (blue) shows a sharp Drude peak at $\omega=0$ due to the presence of long-lived quasiparticles. The PG (green, red, orange) shows a shifted Drude peak at a non-zero frequency which is reminiscent of the mid-infrared (MIR) peak seen experimentally in the cuprates. Inset shows the robustness of the position of the MIR peak through the PG, indicating that it's a universal feature within the PG of our model.}
	\label{optical}
\end{figure}
To further probe the breakdown of coherent transport in the pseudogap (PG) regime, we examine the product of the impurity spectral function $A(\omega)$ and the scattering lifetime $\tau(\omega)$, which approximates the optical conductivity $\sigma(\omega)$ via a Drude-type relation. As shown in the right panel panel of Fig.~\ref{optical}, this quantity exhibits a sharp Drude peak at $\omega = 0$ in the Fermi liquid (FL) regime (blue), consistent with long-lived quasiparticles. In stark contrast, the PG regime (green, red, orange) shows a pronounced suppression of low-frequency conductivity and the emergence of a finite-frequency peak centered near $\omega/D \sim 0.3{-}0.4$ (where $D$ is the kinetic energy bandwidth).

>>>>

%\documentclass{article}
\documentclass[
preprint,
superscriptaddress,
groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
%aps,superscriptaddress,
aps,pra,
%rmp,
%prstab,
%prstper,
%floatfix,
10pt
]{revtex4-2}
%\usepackage{amsmath,amssymb,bm,graphicx,braket}
\usepackage[margin=1.5cm]{geometry}
%\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{graphicx,braket,color,bm,amsmath,amssymb,braket}
% \graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}
\usepackage[]{natbib}

\usepackage{etoolbox}

\makeatletter
\apptocmd{\thebibliography}{\global\c@NAT@ctr 80\relax}{}{}
\makeatother
\begin{document}
\title{Supplementary Materials for ``Mott Criticality as the Confinement Transition of a Pseudogap-Mott Metal"}

\author{Abhirup Mukherjee}
\affiliation{
 Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}
\author{S. R. Hassan}
\affiliation{The Institute of Mathematical Sciences, C.I.T. Campus, Chennai 600 113, India}
\author{Anamitra Mukherjee}
\affiliation{School of Physical Sciences, National Institute of Science, Education and Research, HBNI, Jatni 752050, India}
\author{N. S. Vidhyadhiraja}
\affiliation{Theoretical Sciences Unit, Jawaharlal Nehru Center for Advanced Scientific Research, Jakkur, Bengaluru 560064, India}
\author{A. Taraphder}
\affiliation{Department of Physics, Indian Institute of Technology Kharagpur, Kharagpur 721302, India}
\author{Siddhartha Lal}
\affiliation{
 Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}

@@

The eigenstates \(\ket{\Psi_{s}}\) (\(s = \left({\bf k}, n\right)\)) of the lattice Hamiltonians 
%obtained using eq.~\ref{tilingPrescriptionFinal} also 
enjoy a {\it many-body} Bloch's theorem~[53]
%\cite{stoyanova}, 
because the tiling procedure restores the translation symmetry of the Hamiltonian (as shown in eq.~\ref{translationSymmetry}). This means that the {\it local} eigenstates \(\ket{\psi_n\left({\bf r}_d\right)}\) (with the impurity located at an arbitrary position \({\bf r}_d\)) of the unit cell auxiliary model Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) %defined in eq.~\ref{unitCellHamiltonian} 
can be used to construct eigenstates of the lattice Hamiltonian. The index \(n(=0,1,\ldots)\) in the subscript indicates that it is the \(n^\text{th}\) eigenstate of the auxiliary model.

The state \(\ket{\psi_n\left({\bf r}_d\right)}\) does not specify the position of the zeroth site, because the unit cell Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) itself has been averaged over \(\mathcal{Z}\) zeroth sites. Accordingly, we can express the averaged eigenstate \(\ket{\psi_n\left({\bf r}_d\right)}\) as
\begin{equation}\begin{aligned}
	\ket{\psi_n\left({\bf r}_d\right)} = \frac{1}{\sqrt\mathcal{Z}}\sum_{{\bf z} \in \text{NN}({\bf r}_d)}\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}\) is an auxiliary model eigenstate with the impurity and zeroth sites placed at \({\bf r}_d\) and \({\bf z}\). With this in mind, the following unnormalised combination of the auxiliary model eigenstates satisfies a many-particle equivalent of Bloch's theorem~[53]:
%\cite{stoyanova}:
\begin{equation}\begin{aligned}\label{eigenstateProposal}
	\ket{\Psi_{s}} \equiv \ket{\Psi_{{\bf k}, n}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(N\) is the total number of lattice sites and \({\bf r}_d\) is summed over all lattice spacings. The set of \(n=0\) states form the lowest band in the spectrum of the lattice, while higher values of \(n\) produce the more energetic bands. The ground state \(s = s_0\) is obtained by setting \({\bf k}\) and \(n\) to 0:
\begin{equation}\begin{aligned}\label{groundstateProposal}
	\ket{\Psi_\text{gs}} \equiv \ket{\Psi_{s_0}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d, {\bf z}\right)}
\end{aligned}\end{equation}

@@

\subsection{The unitary renormalisation group method}
In order to obtain the various low-energy phases of our impurity model, we perform a scaling analysis of the associated Hamiltonian using the recently developed unitary renormalisation group (URG) method ~[19,39,40].
%\cite{anirbanurg1,anirbanurg2}. 
The method has been applied successfully on a wide variety of problems of correlated fermions~\cite{santanukagome,1dhubjhep1,anirbanmott1,siddharthacpi,anirban_kondo,Patra_2023}. The method proceeds by resolving quantum fluctuations in high-energy degrees of freedom, leading to a low-energy Hamiltonian with renormalised couplings and new emergent degrees of freedom. Typically, for a system with Fermi energy \(\epsilon_F\) and bandwidth \(E_N\), the sequence of isoenergetic shells \(\left\{E_{(j)}\right\}, E_{j}\in \left[E_0, E_N\right] \) define the states whose quantum fluctuations we sequentially resolve. The momentum states lying on shells \(E_N\) that are far away from the Fermi surface comprise the UV states, while those on shells near the Fermi surface comprise the IR states.

As a result of the URG transformations, the Hamiltonian \(H_{(j)}\) at a given RG step \(j\) involves scattering processes between the \(k-\)states that have energies lower than \(D_{(j+1)}\). The unitary transformation \(U_{(j)}\) is then defined so as to remove the number fluctuations of the currently most energetic set of states \(D_{(j)}\)~[39,40]:
%\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	H_{(j-1)} = U_{(j)} H_{(j)} U^\dagger_{(j)}~, \text{such that} ~\left[H_{(j-1)}, \hat n_{j}\right] =0~.
\end{eqnarray}
The eigenvalue of $\hat{n}_{j}$ has, thus, been rendered an integral of motion (IOM) under the RG transformation.

The unitary transformations can be expressed in terms of a generator \(\eta_{(j)}\) that has fermionic algebra~[39,40]:
%\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	\label{unitary}
	U_{(j)} = \frac{1}{\sqrt 2}\left(1 + \eta_{(j)} - \eta_{(j)}^\dagger\right)~,~ \quad\left\{ \eta_{(j)},\eta_{(j)}^\dagger \right\} = 1~,
\end{eqnarray}
where \(\left\{\cdot\right\}\) is the anticommutator. The unitary operator \(U_{(j)}\) that appears in Eq.~\eqref{unitary} can be cast into the well-known general form \(U = e^\mathcal{S}, \mathcal{S} = \frac{\pi}{4}\left( \eta^\dagger_{(j)} - \eta_{(j)} \right)\) that a unitary operator can take, defined by an anti-Hermitian operator \(\mathcal{S}\). The generator \(\eta_{(j)}\) is given by the expression~[39,40]:
%\cite{anirbanurg1,anirbanurg2}
\begin{eqnarray}
	\eta^\dagger_{(j)} = \frac{1}{\hat \omega_{(j)} - \text{Tr}\left(H_{(j)} \hat n_{j}\right) } c^\dagger_{j} \text{Tr}\left(H_{(j)}c_{j}\right)~.
\end{eqnarray}
The operators \(\eta_{(j)},\eta^\dagger_{(j)}\) behave as the many-particle analogues of the single-particle field operators \(c_j,c^\dagger_j\) - they change the occupation number of the single-particle Fock space \(\ket{n_j}\).  The important operator \(\hat \omega_{(j)}\) originates from the quantum fluctuations that exist in the problem because of the non-commutation of the kinetic energy terms and the interaction terms in the Hamiltonian:
\begin{eqnarray}
	\hat \omega_{(j)} = H_{(j-1)} - H^i_{(j)}~.
	\label{omega}
\end{eqnarray}
\(H^i_{(j)}\) is the part of \(H_{(j)}\) that commutes with \(\hat n_j\) but does {\it not} commute with at least one \(\hat n_l\) for \(l < j\). The RG flow continues up to energy \(D^*\), where a fixed point is reached from the vanishing of the RG function. 
Detailed comparisons of the URG with other methods (e.g., the functional RG, spectrum bifurcation RG etc.) can be found in Refs.~[39,40].
%\cite{anirbanmott1,anirbanurg1}.

@@

\section{Tiling formalism: Mapping from the auxiliary model to the tiled lattice model}
The eigenstates \(\ket{\Psi_{s}}\) (\(s = \left({\bf k}, n\right)\)) of the lattice Hamiltonians obtained using  the tiling procedure outlined in the main text enjoy a {\it many-body} Bloch's theorem~[53],
%\cite{stoyanova}, 
because the tiling procedure restores the translation symmetry of the Hamiltonian. This means that the {\it local} eigenstates \(\ket{\psi_n\left({\bf r}_d\right)}\) (with the impurity located at an arbitrary position \({\bf r}_d\)) of the unit cell auxiliary model Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) can be used to construct eigenstates of the lattice Hamiltonian. The index \(n(=0,1,\ldots)\) in the subscript indicates that it is the \(n^\text{th}\) eigenstate of the auxiliary model.

The state \(\ket{\psi_n\left({\bf r}_d\right)}\) does not specify the position of the zeroth site, because the unit cell Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) itself has been averaged over \(\mathcal{Z}\) zeroth sites. Accordingly, we can express the averaged eigenstate \(\ket{\psi_n\left({\bf r}_d\right)}\) as
\begin{equation}\begin{aligned}
	\ket{\psi_n\left({\bf r}_d\right)} = \frac{1}{\sqrt\mathcal{Z}}\sum_{{\bf z} \in \text{NN}({\bf r}_d)}\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}\) is an auxiliary model eigenstate with the impurity and zeroth sites placed at \({\bf r}_d\) and \({\bf z}\). With this in mind, the following unnormalised combination of the auxiliary model eigenstates satisfies a many-particle equivalent of Bloch's theorem~[53]:
%\cite{stoyanova}:
\begin{equation}\begin{aligned}\label{eigenstateProposal1}
	\ket{\Psi_{s}} \equiv \ket{\Psi_{{\bf k}, n}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(N\) is the total number of lattice sites and \({\bf r}_d\) is summed over all lattice spacings. The set of \(n=0\) states form the lowest band in the spectrum of the lattice, while higher values of \(n\) produce the more energetic bands. The ground state \(s = s_0\) is obtained by setting \({\bf k}\) and \(n\) to 0:
\begin{equation}\begin{aligned}\label{groundstateProposal1}
	\ket{\Psi_\text{gs}} \equiv \ket{\Psi_{s_0}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d, {\bf z}\right)}
\end{aligned}\end{equation}

@@

The ${\bf q}_1={\bf q}_2$ component of the Hamiltonian shows the emergence of the exactly solvable Hatsugai-Kohmoto model~[42,43]
%\cite{Baskaran1991,Hatsugai1992} 
at the critical point. The correlation term $\mathcal{U}\sum_{{\bf q}, \sigma}r_{{\bf q} \sigma}r_{{\bf q} \bar\sigma}$ leads to a transfer of spectral weight across the Fermi surface and separates the available states into three classes:
\begin{equation}\begin{aligned}
\braket{n_{\bf q}} = \begin{cases}
    2, \quad \epsilon_{\bf q} < -|\mathcal{U}/2|~,\\
    1, \quad |\mathcal{U}/2| > \epsilon_{\bf q} > -|\mathcal{U}/2|~,\\
    0, \quad |\mathcal{U}/2| > \epsilon_{\bf q}~.
\end{cases}
\end{aligned}\end{equation}
Different from a Fermi liquid is the emergence of the highly degenerate single-occupied region in the middle, and this gives rise to non-Fermi liquid excitations.

@@

\section{Luttinger’s theorem in the presence of Luttinger surfaces.}
A natural question is whether Luttinger’s theorem continues to hold in our model in the presence of Luttinger surfaces. It has been shown that particle-hole symmetric systems satisfy a generalized version of Luttinger’s theorem~[34],
%\cite{seki2017topological}, 
wherein the Luttinger volume $V_L$ is given by the difference between the number of poles and zeros of the single-particle Green’s function enclosed by the FS. This statement remains valid even when the self-energy $\Sigma(\mathbf{k}, \omega)$ diverges~[32].
%\cite{Phillips2013}. 
In our case, particle-hole symmetry is preserved and the system remains at half-filling, ensuring the total number of occupied states remains constant. Prior to the Lifshitz transition, gapless excitations on the FS contribute one pole per momentum state to the Luttinger count. Inside the PG phase, these gapless $k$-states persist, but the emergence of gapped regions redistributes spectral weight: doubly occupied states on Luttinger surfaces become energetically favorable due to the attractive $W$-interaction. These doubly occupied states are degenerate with the empty states by particle-hole symmetry, and thus, on average, one state remains occupied. The net result is that the number of occupied states - hence the Luttinger volume - remains unchanged. This argument extends to the MI, wherein the FS is entirely replaced by a Luttinger surface with zero quasiparticle weight.

\section{Additional Results on the Fermi Liquid \& Pseudogap-Mott metal Phases}
We present additional figures of the spin (Fig.\ref{spinCorr}) and charge (Fig.\ref{cfnode}) correlations in the impurity model, the impurity spectral function (Fig.\ref{specfunc}), the spin correlations (Fig.\ref{tiledSF}), entanglement entropy (Fig.\ref{tiledEntanglement} (upper panel)) and mutual function (Fig.\ref{tiledEntanglement} (lower panel)) of the tiled model in the pseudogap phase.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.24\textwidth]{SF-1.pdf}
	\includegraphics[width=0.24\textwidth]{SF-2.pdf}
	\includegraphics[width=0.24\textwidth]{SF-3.pdf}
	\includegraphics[width=0.24\textwidth]{SF-4.pdf}
	\caption{\(k-\)space distribution of spin-spin correlation \(\chi_s(d,\vec k)\) between the impurity spin and momentum states in the conduction bath, calculated for a $69\times 69$ $k$-space grid. The last three figures show how \(k-\)points starting from the antinode progressively exit the Kondo cloud, the node being the last \(k-\)point to decouple from the impurity.}
	\label{spinCorr}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.24\textwidth]{cfnode-1.pdf}
	\includegraphics[width=0.24\textwidth]{cfnode-2.pdf}
	\includegraphics[width=0.24\textwidth]{cfantinode-1.pdf}
	\includegraphics[width=0.24\textwidth]{cfantinode-2.pdf}
	\caption{Charge correlation \(\chi_c\) starting from antinode (left) and node (right), at the beginning of the pseudogap phase, calculated for a $49\times 49$ $k$-space grid. Strong node-antinode correlations are clearly visible.}
	\label{cfnode}
\end{figure}
